\chapter{The motivation}

\section{Some problems at large scale}

\begin{itemize}
    \item Proofs in large scale.
    \item Stability of the huge programs.
    \item Correctness of huge programs. 
\end{itemize}

The aim is to know if our code is correct. 

\section{Source of Knowledge}

\mkpt{Knowledge source}
\begin{itemize}
    \item authority
    \item induction of reasoning: using testcases, 
    \item deductive: building proofs for already known knowledge
\end{itemize}

To fully make it useful, we should use deductive reasoning. 

\mkpt{Type systems is crucial to verifying} A compiler can make deductive reasoning. 
This may make programs sound(they won't go wrong). The complete means it can't
be wrong. 

\mkpt{The software crisis} As software grow bigger in size, we find that humans
cannot well control the all the programs underlying the code, especially when
collaborating with others. 

\mkpt{Formal methods}

\mkpt{Degree of formal systems} 

Formal specifications: Only the definition of every object is formally used.
Like all the manuals used in the operating systems.(Popular notation:
Z-notation)

Formal specifications and proofs about programs: Not only do we have the all
the formal specifications, but also proofs about programs are made . 

Formal specs and proofs about programs and machine checking: We have our proofs
checked by machines in all. 

\mkpt{Negation from Maths}Gödel found his first incompleteness theorem: Any arithmetic system that
contains recursive ideas must be incomplete -- that is a there is a theorem that
can neither be proved or disproved.

\mkpt{Rice's Theorem} Every property is undecidable unless a trivial one. (e.g.
it's true for all programs or the property of syntax is true). 

Good part is proofs too detail and mechanically can be handed over to computers. 

\mkpt{Some formal verification systems} TLA+(Temporary Logic Actions) is a 
framework for programs. Coq is another one. 

\section{The Coq System}

\mkpt{Capability} 
\begin{itemize}
    \item Define computable functions and logical predicates
    \item State mathematical theorems theorems and software specifications
    \item Develop formal proofs of theorems interactively 
    \item Machine check these proofs
    \item Extract verified programs to OCaml, Haskell, Scheme. 
\end{itemize}


\mkpt{The way we interact} Cooperating together to generate a proof and to
validate if the proof is true. 
